---
title: "simulation_with_autocorrelation"
output: html_document
---

```{r functions, warning=FALSE, error=FALSE}

library(tidyverse)
library(auriel)

############
#
# Generate twenty populations
#
################

# mu = mean population
# rho = autocorrelation 
# CV = Coefficient of Variation
# spp = number of species

generate_species <- function(nyears=100, mu=1000,rho=0.5,CV=0.2, spp=20) {
  sim.years <- nyears+10
  ln_SD <- sqrt(log(CV^2+1))  #SD parameter of lognormal distribution with desired variability
  ln_SD_e <- sqrt(ln_SD^2/(1-rho^2))  #SD for first value
  time_series_temp <- matrix(rnorm(mean=0,sd=1,n=sim.years*spp),ncol=spp, nrow=sim.years)
  time_series <- matrix(nrow=sim.years,ncol=spp)
  time_series[1,] <- ln_SD_e*time_series_temp[1,]
  for (i in 2:(sim.years)) {
    time_series[i,] <- rho*time_series[(i-1),]+time_series_temp[i,]*ln_SD
  }
  time_series <- time_series + log(mu) - 0.5*ln_SD_e^2
  time_series <- exp(time_series)
  return(time_series[-c(1:10),])
}

################
#
# Grab the two populations with highest pop at time t=1, see what the trend really is
#
################


realtrend <- function(time_series){
  
  ## finds the two species with the highest populations in year 1 
  one <- time_series[,which(time_series[1,]==sort(time_series[1,], TRUE)[1])]
  two <- time_series[,which(time_series[1,]==sort(time_series[1,], TRUE)[2])]
  year = 1:nrow(time_series)
  
  ## puts the three vectors into one data frame, ready to be sampled from  
  highestspecies <-  data.frame(one, two, year)
  
  ## Creates empty object to put results into
  values <- data.frame(beta=c(NA,NA),SE=c(NA,NA),pvalue=c(NA,NA), rsquared=c(NA,NA))
  
  ## run two models, one on the first species, one on the second
  speciesmodel1 <- lm(data=highestspecies, one ~ year)
  speciesmodel2 <- lm(data=highestspecies, two ~ year)
  
  ## extract the beta coefficient, pvalue and rsquared for the first species
  values[1,1] <- speciesmodel1$coefficients[2]
  values[1,2] <- summary(speciesmodel1)$coefficients[,2][2]
  values[1,3] <- summary(speciesmodel1)$coefficients[,4][2]
  values[1,4] <- summary(speciesmodel1)$r.squared
  
  ## extract the beta coefficient, pvalue and rsquared for the second species
  values[2,1] <- speciesmodel2$coefficients[2]
  values[2,2] <- summary(speciesmodel2)$coefficients[,2][2]
  values[2,3] <- summary(speciesmodel2)$coefficients[,4][2]
  values[2,4] <- summary(speciesmodel2)$r.squared
  
  ## puts everything in a list to be output
  realstuff <- list()
  realstuff[["real_values_highest"]] <- values
  realstuff[["real_data_highest"]] <- highestspecies
  
  ###########
  cols <- base::sample(1:ncol(time_series),2)
  one <- time_series[,cols[1]]
  two <- time_series[,cols[2]]
  year = 1:nrow(time_series)
  
  ## puts the three vectors into one data frame, ready to be sampled from  
  randomspecies <-  data.frame(one, two, year)
  
  ## Creates empty object to put results into
  values <- data.frame(beta=c(NA,NA),SE=c(NA,NA),pvalue=c(NA,NA), rsquared=c(NA,NA))
  
  ## run two models, one on the first species, one on the second
  speciesmodel1 <- lm(data=randomspecies, one ~ year)
  speciesmodel2 <- lm(data=randomspecies, two ~ year)
  
  ## extract the beta coefficient, pvalue and rsquared for the first species
  values[1,1] <- speciesmodel1$coefficients[2]
  values[1,2] <- summary(speciesmodel1)$coefficients[,2][2]
  values[1,3] <- summary(speciesmodel1)$coefficients[,4][2]
  values[1,4] <- summary(speciesmodel1)$r.squared
  
  ## extract the beta coefficient, pvalue and rsquared for the second species
  values[2,1] <- speciesmodel2$coefficients[2]
  values[2,2] <- summary(speciesmodel2)$coefficients[,2][2]
  values[2,3] <- summary(speciesmodel2)$coefficients[,4][2]
  values[2,4] <- summary(speciesmodel2)$r.squared
  
  ## puts everything in a list to be output
  realstuff[["real_values_random"]] <- values
  realstuff[["real_data_random"]] <- randomspecies
  
  ########
  ########
  ########
  
  ## finds the two species with the highest populations in year 1 
  ## puts the three vectors into one data frame, ready to be sampled from  
  allspecies <-  data.frame(time_series, year)
  
  ## Creates empty object to put results into
  values <- data.frame(beta=rep(NA,times=20),SE=NA,pvalue=NA, rsquared=NA)
  
  for(column in 1:20){
  ## run two models, one on the first species, one on the second
  speciesmodel1 <- lm(allspecies[,column] ~ allspecies$year)

  ## extract the beta coefficient, pvalue and rsquared for the first species
  values[column,1] <- speciesmodel1$coefficients[2]
  values[column,2] <- summary(speciesmodel1)$coefficients[,2][2]
  values[column,3] <- summary(speciesmodel1)$coefficients[,4][2]
  values[column,4] <- summary(speciesmodel1)$r.squared
}  

  ## puts everything in a list to be output
  realstuff[["real_values_all_pops"]] <- values
  return(realstuff)
}


##################
#
# sample the real population with some frequency of sampling effort with error
#
##################

samplingeffortwitherror <- function(time_series, freq=NA, sd=0.1){
  
  ## The default is to have freq=NA, which means that the function 
  ## will randomly pull a random set of years to sample on
  ## This checks to see if freq=NA, and if so, does that
  ## If we set freq = to some number it will bipass that and use that instead
  if(is.na(freq[1])){
    freq <- sort(base::sample(1:100, sample(1:60, 1), replace=FALSE))
  }
  
  ## finds the two species with the highest populations in year 1 
  one <- time_series[,which(time_series[1,]==sort(time_series[1,], TRUE)[1])]
  two <- time_series[,which(time_series[1,]==sort(time_series[1,], TRUE)[2])]
  year = 1:nrow(time_series)
  
  ## puts the three vectors into one data frame, ready to be sampled from  
  twospecies <-  data.frame(one, two, year)
  
  ## If we have set freq = to some single value, this if statement runs
  if(length(freq)==1){
    sampled_data <- twospecies[seq(from=1, to=nrow(time_series), by=freq),]
  }
  
  ## if freq is a vector of years, than this one runs
  if(length(freq)>1){
    sampled_data <- twospecies[freq,]
  }
  
  ## This takes the years we just sampled, and applies an error term
  ## since its pretty hard to actually estimate the population
  sdata <- list()
  sampled_data[,1:2] <-  sampled_data[,1:2] * exp(rnorm(1, mean=0, sd=sd)-0.5*sd^2)
  sdata[["sampled_data_highest"]] <- sampled_data
  
  ####
  #
  # two randomly selected
  #
  ####
  
  ## The default is to have freq=NA, which means that the function 
  ## will randomly pull a random set of years to sample on
  ## This checks to see if freq=NA, and if so, does that
  ## If we set freq = to some number it will bipass that and use that instead
  if(is.na(freq[1])){
    freq <- sort(base::sample(1:100, sample(1:60, 1), replace=FALSE))
  }
  
  ## grabs two species at random
  one <- time_series[,base::sample(1:ncol(time_series),1)]
  two <- time_series[,base::sample(1:ncol(time_series),1)]
  year = 1:nrow(time_series)
  
  ## puts the three vectors into one data frame, ready to be sampled from  
  twospecies <-  data.frame(one, two, year)
  
  ## If we have set freq = to some single value, this if statement runs
  if(length(freq)==1){
    sampled_data <- twospecies[seq(from=1, to=nrow(time_series), by=freq),]
  }
  
  ## if freq is a vector of years, than this one runs
  if(length(freq)>1){
    sampled_data <- twospecies[freq,]
  }
  
  ## This takes the years we just sampled, and applies an error term
  ## since its pretty hard to actually estimate the population
  sampled_data[,1:2] <-  sampled_data[,1:2] * exp(rnorm(1, mean=0, sd=sd)-0.5*sd^2)
  sdata[["sampled_data_random"]] <- sampled_data
  
  return(sdata)
}

##################
#
#  What trend do we see in the sampled data? 
#
##################

sampletrend <- function(sampled_data){

  ##
  # Two Highest
  ##
    
  ## Create empty object to store results in
  sampled_values <- data.frame(beta=c(NA,NA),SE=c(NA,NA),pvalue=c(NA,NA),rsquared=c(NA,NA))
  
  ## run two models, one on the first species, one on the second
  speciesmodel1 <- lm(data=sampled_data$sampled_data_highest, one ~ year)
  speciesmodel2 <- lm(data=sampled_data$sampled_data_highest, two ~ year)
  
  ## extract the beta coefficient, pvalue and rsquared for the first species
  sampled_values[1,1] <- speciesmodel1$coefficients[2]
  sampled_values[1,2] <- summary(speciesmodel1)$coefficients[,2][2]
  sampled_values[1,3] <- summary(speciesmodel1)$coefficients[,4][2]
  sampled_values[1,4] <- summary(speciesmodel1)$r.squared
  
  ## extract the beta coefficient, pvalue and rsquared for the second species
  sampled_values[2,1] <- speciesmodel2$coefficients[2]
  sampled_values[2,2] <- summary(speciesmodel2)$coefficients[,2][2]
  sampled_values[2,3] <- summary(speciesmodel2)$coefficients[,4][2]
  sampled_values[2,4] <- summary(speciesmodel2)$r.squared
  
  sampled_trend <- list()
  sampled_trend[["sampled_trend_highest"]] <- sampled_values
  ##
  # Two Random
  ## 
  
  ## Create empty object to store results in
  sampled_values <- data.frame(beta=c(NA,NA),SE=c(NA,NA),pvalue=c(NA,NA),rsquared=c(NA,NA))
  
  ## run two models, one on the first species, one on the second
  speciesmodel1 <- lm(data=sampled_data$sampled_data_random, one ~ year)
  speciesmodel2 <- lm(data=sampled_data$sampled_data_random, two ~ year)
  
  ## extract the beta coefficient, pvalue and rsquared for the first species
  sampled_values[1,1] <- speciesmodel1$coefficients[2]
  sampled_values[1,2] <- summary(speciesmodel1)$coefficients[,2][2]
  sampled_values[1,3] <- summary(speciesmodel1)$coefficients[,4][2]
  sampled_values[1,4] <- summary(speciesmodel1)$r.squared
  
  ## extract the beta coefficient, pvalue and rsquared for the second species
  sampled_values[2,1] <- speciesmodel2$coefficients[2]
  sampled_values[2,2] <- summary(speciesmodel2)$coefficients[,2][2]
  sampled_values[2,3] <- summary(speciesmodel2)$coefficients[,4][2]
  sampled_values[2,4] <- summary(speciesmodel2)$r.squared
  
  sampled_trend[["sampled_trend_random"]] <- sampled_values
  
  return(sampled_trend)
}

##################
#
#  One simulation, all in one simple function
#
##################

onesimulation <- function(freq=NA){
  
  ## Generate the time series for 20 species
  time_series <- generate_species()
  
  ## Figure out the 'real trend'
  rdat <- realtrend(time_series)
  
  ## Sample the real population data, with some error
  sampled_data <- samplingeffortwitherror(time_series, freq=freq, sd=0.1)
  
  ## figure out what the trend is for the sampled_data
  svalues <- sampletrend(sampled_data)
  
  ## output everything into one master list
  outputs <- list()
  outputs[["twenty_species"]] <- time_series
  outputs[["real_values_highest"]] <- rdat$real_values_highest
  outputs[["real_values_random"]] <- rdat$real_values_random
  outputs[["real_data_highest"]] <- rdat$real_data_highest
  outputs[["real_data_random"]] <- rdat$real_data_random
  outputs[["real_values_all_pops"]] <- rdat$real_values_all_pops
  outputs[["sampled_data_highest"]] <- sampled_data$sampled_data_highest
  outputs[["sampled_data_random"]] <- sampled_data$sampled_data_random
  outputs[["sampled_values_highest"]] <- svalues$sampled_trend_highest
  outputs[["sampled_values_random"]] <- svalues$sampled_trend_random
  return(outputs)
}

####################
#
#  Run the simulation many times, store everything in a list and spit it out
#
####################

manysimulations <- function(sims=1000, freq=NA){
  
  ## Create empty objects to store results in
  ## sampledata is a list because the number of rows could vary from sim to sim
  real_values_highest <- data.frame(beta=rep(NA,sims*2),SE=rep(NA,sims*2),pvalue=rep(NA,sims*2),rsquared=rep(NA,sims*2))
  real_values_random <- data.frame(beta=rep(NA,sims*2),SE=rep(NA,sims*2),pvalue=rep(NA,sims*2),rsquared=rep(NA,sims*2))
  real_values_all_pops <- list()
  sampled_values_highest <- data.frame(beta=rep(NA,sims*2),SE=rep(NA,sims*2),pvalue=rep(NA,sims*2),rsquared=rep(NA,sims*2))
  sampled_values_random <- data.frame(beta=rep(NA,sims*2),SE=rep(NA,sims*2),pvalue=rep(NA,sims*2),rsquared=rep(NA,sims*2))
  real_data_highest <- as.data.frame(matrix(ncol=(2*sims), nrow=100))
  real_data_random <- as.data.frame(matrix(ncol=(2*sims), nrow=100))
  sampled_data_highest <- list()
  sampled_data_random <- list()
  twenty_species <- list()
  
  ## Runs through the onesimulation function sims number of times
  ## stores the various results in the different objects
  for(i in seq(1,(sims*2),by=2)){
    dd <- onesimulation(freq=freq)
    real_values_highest[i:(i+1),] <- dd$real_values_highest
    real_values_random[i:(i+1),] <- dd$real_values_random
    real_values_all_pops[[i]] <- dd$real_values_all_pops
    sampled_values_highest[i:(i+1),] <- dd$sampled_values_highest
    sampled_values_random[i:(i+1),] <- dd$sampled_values_random
    real_data_highest[,i:(i+1)] <- dd$real_data_highest[,1:2]
    real_data_random[,i:(i+1)] <- dd$real_data_random[,1:2]
    sampled_data_highest[[i]] <- dd$sampled_data_highest[,1:2]
    sampled_data_random[[i]] <- dd$sampled_data_random[,1:2]
    colnames(dd$twenty_species) <- paste0("spp",1:20)
    t_species <- data.frame(dd$twenty_species) %>% mutate(year=1:100) %>% gather("variable","value", -year) %>% mutate(sim=i)
    twenty_species[[i]] <- t_species
  }
  
  ## load everything up into one master list to be outputted
  outputs <- list()
  outputs[["real_values_highest"]] <- real_values_highest
  outputs[["real_values_random"]] <- real_values_random
  outputs[["real_values_all_pops"]] <- do.call(rbind, real_values_all_pops)
  outputs[["sampled_values_highest"]] <- sampled_values_highest
  outputs[["sampled_values_random"]] <- sampled_values_random
  outputs[["real_data_highest"]] <- real_data_highest
  outputs[["real_data_random"]] <- real_data_random
  outputs[["sampled_data_highest"]] <- sampled_data_highest 
  outputs[["sampled_data_random"]] <- sampled_data_random
  outputs[["twenty_species"]] <- twenty_species
  
  return(outputs)
  
}

```


Bursts of sampling with bursts of nothing (perhaps representative of a project that is picked up, dropped, picked up again)

```{r bursts of sampling}
# bursts_of_sampling <- c(1,2,3,4,5,6,15,16,18,19,37,38,39,47,48,56,69,70,71,72,73,74,75,76,77,78,91,92,93,97,99)
# 
# dat <- manysimulations(freq=bursts_of_sampling, sims=100000)
# 
# save(dat, file="~/negative_population_trends/bursts_of_sampling.Rdata")

load("../../Dropbox/negative_population_trends/bursts_of_sampling.Rdata")

# the highest two populations at year 1 

values <- dat$real_values_highest
svalues <- dat$sampled_values_highest

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")


sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

# two randomly selected species

values <- dat$real_values_random
svalues <- dat$sampled_values_random

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")


sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

```

# Sampling Every 2 Years

```{r every two years}

# dat <- manysimulations(freq=2, sims=100000)
# 
# 
# save(dat, file="~/negative_population_trends/every_two_years.Rdata")

load("../../Dropbox/negative_population_trends/every_two_years.Rdata")

# the highest two populations at year 1

values <- dat$real_values_highest
svalues <- dat$sampled_values_highest

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")

sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

# two randomly selected species

values <- dat$real_values_random
svalues <- dat$sampled_values_random

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")


sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

```


# Sampling Every 5 Years

```{r every five years}

# dat <- manysimulations(freq=5, sims=100000)
# 
# save(dat, file="~/negative_population_trends/every_five_years.Rdata")

# the highest two populations at year 1

load("../../Dropbox/negative_population_trends/every_five_years.Rdata")

values <- dat$real_values_highest
svalues <- dat$sampled_values_highest

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05 ) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05 ) %>% mutate(type="surveyeddata_significant")



sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

# two randomly selected species

values <- dat$real_values_random
svalues <- dat$sampled_values_random

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05 ) %>% mutate(type="surveyeddata_significant")


sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

```


# Random Sampling

```{r random sampling}

# dat <- manysimulations(sims=100000)
# 
# 
# save(dat, file="~/negative_population_trends/random.Rdata")

# the highest two populations at year 1

load("../../Dropbox/negative_population_trends/random.Rdata")

values <- dat$real_values_highest
svalues <- dat$sampled_values_highest

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05 ) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")


sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

# two randomly selected species

values <- dat$real_values_random
svalues <- dat$sampled_values_random

negvaluessig <- values %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="realtrend_significant")
negsvaluessig <- svalues %>% filter(pvalue<=0.05 & beta < 0) %>% mutate(type="surveyeddata_significant")

nrow(negsvaluessig)/nrow(svalues) # proportion of the sampled population that has significant negative trend
nrow(negvaluessig)/nrow(values) # proportion of the real population that has signficaint negative trend

valuessig <- values %>% filter(pvalue<=0.05) %>% mutate(type="realtrend_significant")
svaluessig <- svalues %>% filter(pvalue<=0.05) %>% mutate(type="surveyeddata_significant")

sigdat <- bind_rows(valuessig, svaluessig)

model <- aov(data=sigdat, beta ~ type)

summary(model)

ggplot(data=sigdat, aes(y=beta, x=type))+ geom_boxplot() + ylim(-10,10)

```

# Graph Comparing Mean Trend of Randomly Surveyed Populations to Mean Trend of All 20 Populations


```{r graph of mean trends}

mean_survey <- dat$real_values_highest %>% filter(pvalue<=0.05) %>% summarize(mean = mean(beta))

mean_pop <- dat$real_values_all_pops %>% filter(pvalue<=0.05) %>% summarize(mean = mean(beta))

CI_survey <- dat$real_values_highest %>% filter(pvalue<0.05) %>% summarize(lower=quantile(beta, probs=c(0.05)), upper=quantile(beta, probs=0.95))

CI_pop <- dat$real_values_all_pops %>% filter(pvalue<0.05) %>% summarize(lower=quantile(beta, probs=c(0.05)), upper=quantile(beta, probs=0.95))

datdat <- data.frame(rbind(cbind(mean_survey, CI_survey),cbind(mean_pop, CI_pop)))
datdat$type <- c("survey","all_pops")


ggplot(data=datdat)+
  geom_point(aes(x=type, y=mean))+
  geom_errorbar(aes(x=type, ymin=lower, ymax=upper))+
  geom_hline(aes(yintercept=0))

```

# The Proportion of significant declining trends in surveyed population versus all 20 populations

```{r proportions}

(dat$real_values_highest %>% filter(pvalue<0.05 & beta<0) %>% nrow())/(dat$real_values_all_pops %>% filter(pvalue<0.05 & beta<0) %>% nrow())

```



# Figuring showing the two surveyed population's trends overlayed with the overall trend from all 20 populations together, for five seperate simulations. 

```{r trend overlay graph, fig.width=10, fig.height=10}
set.seed(2017) # so the figure looks the same every time

dat <- manysimulations(sims=5)

t20 <- do.call(rbind, dat$twenty_species) %>% # bringing together a data frame of all 20 species
            mutate(sim=rep(c(1,2,3,4,5), each=2000)) # makes a sim column so we can tell which data came from which simulation

dd20 <- lm(data=t20, value ~ year) # simple linear regression of all 20 species

t20$fitted_value <- dd20$fitted.values # adding the fitted values to the data frame

t20mean <- t20 %>% # data frame with data from the 20 populations
            group_by(year) %>% # group them by year so we can calculate year specific Confidence Intervals
            summarize(mean=mean(fitted_value), # fine the mean
                      standD=sd(value), # find the standard deviation
                   sample_size=n(), # sample size 
                   CI=qnorm(0.975)*standD/sqrt(sample_size)) # confidence interval for that particular year

t2 <- dat$real_data_highest %>% # data frame of just the two species (highest populations at year 1) surveyed
          gather("variable","value") %>%  # turning it from wide to long data
          mutate(year=rep(c(1:100),times=10), # creating a year column
                 sim=rep(c(1,2,3,4,5), each=200)) # creating a column indicating what simulation data are from



dd2 <- lm(data=t2, value ~ year) # simple linear regression of just the two surveyed species

t2$fitted_value <- dd2$fitted.values # adding the fitted values to the data frame


ggplot()+
  geom_point(data=t20, col="grey",aes(x=year, y=value, group=variable))+ # all 20 populations background points
  geom_ribbon(data=t20mean, aes(x=year, ymin=mean-CI, ymax=mean+CI), alpha=0.7)+ # draws the confidence invertal for that particular year
  geom_line(data=t20mean, aes(x=year, y=mean))+ # draws the trend line for that particular year
  geom_point(data=t2, aes(x=year, y=value))+ # draws the points from the two sampled populations
  stat_smooth(data=t2, aes(x=year, y=value, group=variable), method="lm")+ # draws the trend line with standard error for each species individually
  theme_krementz()+ # applies a ggplot two theme, named after my advisor David Krementz
  facet_wrap(~sim) # gives each simulation its own facet



```
